Data Anonymization for Open Science: A Case Study

### 2025-05-14 submission_materials/article.tex
added citations of original study where appropriate

\hl{Although the data quality and utility was previously thoroughly explored, previous research on data privacy risks associated with non-traditional anonymization methods and synthetic data generation is rather sparse [BAYREM]. Few frameworks exist that can estimate residual privacy risks on anonymized and synthetic data alike. One of such framework is the Anonymeter framework by Giomi et al \mycite{20}, originally proposed for large, synthetic datasets to evaluate risk of linkage, attribute inference and record singling-out. 
}

@@

The following sections discuss the similarities and differences between the anonymized and original data, and comment on the extent to which the conclusions of the original paper still hold given the anonymized data.

@@

\subsection*{Commute modes and distances (Base-tables 1 and 2 from the original paper)}

@@

Among the three anonymization methods, SynDiffix most closely reproduced original counts and percentages for cross-tabulation of commuting modes (Table 2), with only one deviating record in the cross table. It is closely followed by ARX which showed up to 18 deviations, while SDV provides large differences of up to 179 different commute mode counts in the cross table. Nevertheless, the basic description of these results in original paper, i.e. that active commuting is more frequent in the direction from school to home than in other direction, still holds, but this may be a coincidence in case of SDV as it substantially over- or underestimates the frequencies for both modes of active commuting (walking and wheels). Note also that the anonymized table of ARX contains 1.4\% fewer rows (703 versus 713) due to suppression of rare records.

Similar results were found for commuting distance (Table~3). Here also SDV performs badly, hugely over- or underestimating the distance in most of the table cells, both in central tendency (median) and variability (IQR) of the data. SynDiffix and ARX perform much better, but still with large errors in some of the table cells, especially in cells with low frequencies, e.g. SynDiffix substantially underestimates the median distance of car commuting from school to home (3615$\rightarrow$2584) and ARX overestimates the distance of wheels commuting in same direction (1444$\rightarrow$2235). The verbal description of this results from original paper, i.e. that active commuting groups (Walk, Wheels) typically live
close to school, still holds in case of SynDiffix and ARX, but not in case of SDV.

@@

      \bottomrule
      \end{tabular}
      \end{small}
      \caption{Base-table 2 from the original paper showing the counts and distances in meters (median and IQR) for the original data and the three anonymization methods. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. The bold/italic fonts for counts (N) are as described for Table 2. Distance and IRQ are \textbf{bold font} when the relative error is greater than 30\%, and \textit{italics} when the relative error is greater than 15\%.  Note that the original distances median and IQR don't perfectly match those of the base-table 2 because of differences in the way median and IQR were calculated (Python versus R).}
      \label{tab:table2}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

      \bottomrule
      {\footnotesize * p $\leq$ 0.05, \quad** p $\leq$ 0.01, \quad*** p $\leq$ 0.001}
      \end{tabular}
      \end{small}
      \caption{Part 2 (of 2) of the original paper's Table 3 showing the parameters (regression coefficients) of the linear model for prediction of VO$_2$max by group and distance. \textbf{Bold font} indcates that the anonymized entry is non-significant where the original data is significant or vice versa. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. 
      }
      \label{tab:table3b}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

\subsection*{Statistical significance of regression coefficients (Base-table 3 from the original paper)}

@@

\subsection*{Stratification of children's cardiorespiratory fitness (Base-figure 1 from the original paper)}

The plot for base-figure 1 is given in this paper's Figure 3, which replicates the plot from the original paper as well as gives the corresponding plots for the three anonymized datasets.
In male children, point predictions and prediction intervals in the original paper are quite closely matched with the ones from ARX and SynDiffix, but not with SDV. 
In female children, original statistics are most closely matched by ARX, followed SynDiffix and then by SDV.
SDV gives similar results for male and female children, although they are clearly separated in the original data. SDV performs worst also in estimating prediction interval widths, which are very similar for both sexes, although being quite different in the original plot. ARX and SynDiffix produce interval widths that are more similar to original, but still differ in some cases, e.g. being too narrow in case of walk commuting in ARX and too narrow in females' wheels commuting from school to home in both ARX and SynDiffix.

@@

\section*{Discussion, \hl{conclusions, and future work}}

@@

Recreating the three tables and one plot from the original paper required seven different tables from the blob; 2 with 1 column, 3 with 2 columns, and 2 with 6 columns and a target column (VO$_2$max). In the scripts used for generating the tables and plot, an additional API call to \texttt{SyndiffixBlobReader} was required prior to each analytic task:

@@

Absolute error of the three anonymization methods for the counts and distances in Tables 1  and 2. Each box plot data point is taken from a single cell in Tables 1 and 2.

@@

Normalized error for coefficients in Tables~3 and 4, and Fit (V0$_2$max predicted at mode for median distance) for Figure 3. Normalized error $E$ is computed as $E = \frac{\left| \text{O} - \text{A} \right|}{\max\left( \left| \text{O} \right|, \left| \text{A} \right| \right)}$, where $O$ is the original value and $A$ is the anonymized value. The middle plot displays normalized error only for the datapoints in Tables~3 and 4 where the original coefficient is significant.

@@

Comparison of the VO$_2$max data plots corresponding to the base-figure from the original study.

>>>>

\hl{Although the data quality and utility was previously thoroughly explored, previous research on data privacy risks associated with non-traditional anonymization methods and synthetic data generation is rather sparse [BAYREM]. Few frameworks exist that can estimate residual privacy risks on anonymized and synthetic data alike. One of such framework is the Anonymeter framework by Giomi et al} \mycite{20} \hl{, originally proposed for large, synthetic datasets to evaluate risk of linkage, attribute inference and record singling-out.}

@@

The following sections discuss the similarities and differences between the anonymized and original data, and comment on the extent to which the conclusions of the original paper \mycite{6} still hold given the anonymized data.

@@

\subsection*{Reproduction of commute modes and distances}

@@

Among the three anonymization methods, SynDiffix most closely reproduced original counts and percentages for cross-tabulation of commuting modes (Table 2), with only one deviating record in the cross table. It is closely followed by ARX which showed up to 18 deviations, while SDV provides large differences of up to 179 different commute mode counts in the cross table. Nevertheless, the basic description of these results in original paper \mycite{6}, i.e. that active commuting is more frequent in the direction from school to home than in other direction, still holds, but this may be a coincidence in case of SDV as it substantially over- or underestimates the frequencies for both modes of active commuting (walking and wheels). Note also that the anonymized table of ARX contains 1.4\% fewer rows (703 versus 713) due to suppression of rare records.

Similar results were found for commuting distance (Table~3). Here also SDV performs badly, hugely over- or underestimating the distance in most of the table cells, both in central tendency (median) and variability (IQR) of the data. SynDiffix and ARX perform much better, but still with large errors in some of the table cells, especially in cells with low frequencies, e.g. SynDiffix substantially underestimates the median distance of car commuting from school to home (3615$\rightarrow$2584) and ARX overestimates the distance of wheels commuting in same direction (1444$\rightarrow$2235). The verbal description of this results from original paper \mycite{6}, i.e. that active commuting groups (Walk, Wheels) typically live
close to school, still holds in case of SynDiffix and ARX, but not in case of SDV.

@@

      \bottomrule
      \end{tabular}
      \end{small}
      \caption{Base-table 2 from the original paper showing the counts and distances in meters (median and IQR) for the original data and the three anonymization methods \mycite{6}. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. The bold/italic fonts for counts (N) are as described for Table 2. Distance and IRQ are \textbf{bold font} when the relative error is greater than 30\%, and \textit{italics} when the relative error is greater than 15\%.  Note that the original distances median and IQR don't perfectly match those of the base-table 2 because of differences in the way median and IQR were calculated (Python versus R).}
      \label{tab:table2}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

      \bottomrule
      {\footnotesize * p $\leq$ 0.05, \quad** p $\leq$ 0.01, \quad*** p $\leq$ 0.001}
      \end{tabular}
      \end{small}
      \caption{Part 2 (of 2) of the original paper's Table 3 showing the parameters (regression coefficients) of the linear model for prediction of VO$_2$max by group and distance \mycite{6}. \textbf{Bold font} indcates that the anonymized entry is non-significant where the original data is significant or vice versa. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. 
      }
      \label{tab:table3b}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

\subsection*{Reproduction of statistical significance of regression coefficients}

@@

\subsection*{Reproduction of the stratification of children's cardiorespiratory fitness}

The plot for base-figure 1 is given in this paper's Figure 3, which replicates the plot from the original paper as well as gives the corresponding plots for the three anonymized datasets \mycite{6}.
In male children, point predictions and prediction intervals in the original paper \mycite{6} are quite closely matched with the ones from ARX and SynDiffix, but not with SDV. 
In female children, original statistics are most closely matched by ARX, followed SynDiffix and then by SDV.
SDV gives similar results for male and female children, although they are clearly separated in the original data. SDV performs worst also in estimating prediction interval widths, which are very similar for both sexes, although being quite different in the original plot. ARX and SynDiffix produce interval widths that are more similar to original, but still differ in some cases, e.g. being too narrow in case of walk commuting in ARX and too narrow in females' wheels commuting from school to home in both ARX and SynDiffix.

@@

\section*{Discussion}

@@

Recreating the three tables and one plot from the original paper \mycite{6} required seven different tables from the blob; 2 with 1 column, 3 with 2 columns, and 2 with 6 columns and a target column (VO$_2$max). In the scripts used for generating the tables and plot, an additional API call to \texttt{SyndiffixBlobReader} was required prior to each analytic task:

@@

Absolute error of the three anonymization methods for the counts and distances in Tables 1  and 2. Each box plot data point is taken from a single cell in Tables 2 and 3.

@@

Normalized error for coefficients in Tables~3 and 4, and Fit (V0$_2$max predicted at mode for median distance) for Figure 3. Normalized error $E$ is computed as $E = \frac{\left| \text{O} - \text{A} \right|}{\max\left( \left| \text{O} \right|, \left| \text{A} \right| \right)}$, where $O$ is the original value and $A$ is the anonymized value. The middle plot displays normalized error only for the datapoints in Tables~4 and 5 where the original coefficient is significant.

@@

Comparison of the VO$_2$max data plots corresponding to the base-figure from the original study \mycite{6}.
